---
title: "R Notebook"
output: html_notebook
---

Read the file
```{r}
library(tidyverse)
library(lubridate)
library(ggtext)
library(scales)
spotify_songs_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
```
See the quantity of featuring songs
most of the songs contain feat. as the featuring word
```{r}
spotify_songs_raw %>% 
  filter(str_detect(track_name,"featuring|Featuring"))

```

Performing regex for extract the featuring artist list, and then splitting by ","/"&"/"and" words, to get a vector with the featuring artists for each song.
```{r}
spotify_songs <- spotify_songs_raw %>% 
    mutate(feat_raw = str_extract(track_name,"(?<=(feat\\. )).+?(?=\\)|\\])")) %>% 
  mutate(feat_ls = str_split(feat_raw,", | & |and "),
    num_feat = map_dbl(feat_ls,NROW),
    year = year(ymd(track_album_release_date)))

songs_feat <- spotify_songs %>% 
  filter(str_detect(track_name,"feat.")) %>% 
  filter(!is.na(feat_raw))
```

The song with biggest ammount of featuring artists
```{r}
songs_feat %>% 
  filter(num_feat == max(num_feat)) %>% 
  select(track_name,feat_raw,num_feat)
```
Make a long list with each individual featuring artist
```{r}
songs_feat_long <- songs_feat %>% 
  unnest(feat_ls) %>% 
  rename(feat_artist = feat_ls)

```
Which artist is in most featurings
```{r}
songs_feat_long %>% 
  group_by(feat_artist) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```
Most songs in the list are from 2019
```{r}
ggplot(songs_feat %>% group_by(year) %>% summarize(n = n())) +
  geom_col(mapping = aes(year,n))
```

```{r}
songs_feat_long %>% filter(feat_artist == "Drake")
```
Featurings by genre

```{r}
percent_feat_genre <- spotify_songs %>% 
  group_by(playlist_genre) %>% 
  summarize(n = n(),
    n_feat = sum(!is.na(feat_raw))) %>% 
  mutate(percent_feat = n_feat/n) %>% 
  arrange(desc(percent_feat))
percent_feat_genre
```

Most featured artist by genre?
```{r}
top_feat_by_genre <- songs_feat_long %>% 
  group_by(playlist_genre,feat_artist) %>%
  summarize(n = n()) %>% 
  ungroup() %>% 
  group_by(playlist_genre) %>% 
  filter(n == max(n)) %>% 
   summarize(top_feat_artists = paste(paste(feat_artist,collapse = ", ")," (",unique(n),")"),
     n_top_feat = unique(n))
top_feat_by_genre
```

```{r}
genre_feat_summary <- percent_feat_genre %>% 
  left_join(top_feat_by_genre,by = c("playlist_genre")) %>% 
  mutate(fct_reorder(playlist_genre,percent_feat))
genre_feat_summary
```

```{r}
ggplot(genre_feat_summary) +
  geom_col(mapping = aes(fct_reorder(playlist_genre,percent_feat),n),fill = "grey90",width = 0.5) +
  geom_col(mapping = aes(playlist_genre,n_feat,fill = "Contains feat."),width = 0.5) +
  geom_col(mapping = aes(playlist_genre,n_top_feat,fill = "Top featuring artists")) +
  geom_text(mapping = aes(playlist_genre,n_top_feat,label = top_feat_artists,vjust = -1.5,hjust = "left"),
    color = "#00aeb9",size = 3) +
  geom_text(mapping = aes(playlist_genre,n_feat + 325,label = percent(percent_feat,accuracy = 0.1),hjust = "right"),
    color = "#E87722",size = 3) +
  scale_fill_manual(values = c("#E87722","#00aeb9")) +
  coord_flip() +
  theme_minimal() +
   theme(legend.position = "bottom",
    plot.title = element_markdown(size = 20, hjust = 0.5,color = "#222222", face = "bold"),
    axis.ticks = element_blank(),
    legend.text.align = 0,
     panel.grid.major = element_blank(), 
     panel.grid.minor = element_blank(),
    panel.background = element_blank()) +
  labs(x = "",
    y = "songs",
    fill = "",
    legend = "",
    title = "Rock bands don't like <span style='color:#E87722'>feat</span>uring artists",
    subtitle = "",
    caption = "@perspectivalean")


 ggsave("output/feat_picture.png",plot = last_plot(),width = 31.75/1.5,height = 17.85/1.5,units = c("cm"),dpi = 900)
```

